A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-04-23 tidyindex.qmd
quote

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [XXX tidyr, dplyr, tidymodels, ..., tsibble, cubble] that turn ideas into software quickly and faithfully, as envisioned by @chambers. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->.

>>>>

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [XXX tidyr, dplyr, tidymodels, ..., tsibble, cubble] that "turn ideas into software quickly and faithfully", as envisioned by @chambers. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->.
